---
title: "Psyllid microbiome - Statistics"
author: "Alexander Piper"
date: "07/04/2020"
output:
  html_document:
    highlighter: null
    theme: "flatly"
    code_download: true
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    df_print: paged
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# Knitr global setup - change eval to true to run code
library(knitr)
knitr::opts_chunk$set(echo = TRUE, eval=FALSE, message=FALSE, error=FALSE,fig.show = "hold", fig.keep = "all")
opts_chunk$set(dev = 'png')
```

# Load packages 

```{r install & Load packages} 
#Set required packages
.cran_packages <- c("tidyverse", 
                    "patchwork", 
                    "vegan", 
                    "seqinr",
                    "ape", 
                    "sp",
                    "data.table", 
                    "RColorBrewer",
                    "ggtree", 
                    "castor", 
                    "picante", 
                    "MCMCglmm", 
                    "phylosignal", 
                    "adephylo",
                    "TreeDist",
                    "paco",
                    "phytools")
.bioc_packages <- c("dada2",
                    "phyloseq", 
                    "DECIPHER",
                    "Biostrings",
                    "ShortRead", 
                    "philr",
                    "ALDEx2")

.inst <- .cran_packages %in% installed.packages()
if(any(!.inst)) {
   install.packages(.cran_packages[!.inst])
}
.inst <- .bioc_packages %in% installed.packages()
if(any(!.inst)) {
  if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
  BiocManager::install(.bioc_packages[!.inst], ask = F)
}

sapply(c(.cran_packages,.bioc_packages), require, character.only = TRUE)

# Github packages
devtools::install_github("alexpiper/taxreturn")
library(taxreturn)
devtools::install_github("alexpiper/seqateurs")
library(seqateurs)
devtools::install_github("mikemc/speedyseq")
library(speedyseq)
devtools::install_github('ggloor/CoDaSeq/CoDaSeq')
library(CoDaSeq)
devtools::install_github("DanielSprockett/reltools")
library(reltools)

#Source internal functions
source('R/BDTT.R')
source('R/phylosymbiosis.R')
source('R/helper_functions.R')
options(stringsAsFactors = FALSE)
```

# Make Phyloseq object

Following taxonomic assignment, the sequence table and taxonomic table are merged into a single phyloseq object alongside the sample info csv.

```{r create PS, eval = FALSE}
seqtab_final <- readRDS("output/rds/seqtab_final.rds")
tax <- readRDS("output/rds/tax_IdTaxaExact.rds") 
fitGTR <- readRDS("output/rds/phytree.rds") 

#Load sample information (Put all collection and geographic distances here?)
## ---- samdat ----
samdf <- read_csv("sample_data/Sample_info.csv")  %>%
  dplyr::rename_all(funs( stringr::str_replace_all(., '\\ ', '_')) ) %>%
  mutate(psyllid_spp = psyllid_spp %>% str_remove("\\.") %>%
           str_replace_all(" ", "_") %>%
           str_replace_all("Casuarinicola sp", "Triozid sp")) %>%
  mutate(hostplant_spp = hostplant_spp %>% str_remove("\\.") %>%
           str_replace_all(" ", "_"))

write_csv(samdf, "sample_data/Sample_info2.csv")

# Flag replicated samples
samdf <- samdf %>%
  dplyr::mutate(replicated = Sample_Name %in% (samdf %>% group_by(Sample_Name) %>% 
                                                 add_count() %>%
                                           filter(n >1) %>% 
                                           pull(Sample_Name))) %>%
  separate(Collection, into= c("Lat", "Long"), sep=" ", remove = TRUE) %>%
  separate(Lat, paste("lat",c("d","m","s"), sep="_")) %>%
  separate(Long, paste("long",c("d","m","s"), sep="_" )) %>%
  mutate_at(vars(starts_with("lat")), .funs=as.numeric) %>%
  mutate_at(vars(starts_with("long")), .funs=as.numeric) %>%
  mutate(lat=-(lat_d + lat_m/60 + lat_s/60^2),
            long=long_d + long_m/60 + long_s/60^2) %>%
  dplyr::select(c("SampleID", "Sample_Name", "seqrun", "psyllid_spp", "psyllid_genus", "psyllid_family", "hostplant_spp","Collection_Date","replicated", "lat", "long")) %>%
  as.data.frame(stringsAsFactors=FALSE) %>%
  magrittr::set_rownames(.$SampleID) #Collection

## ---- phyloseq ----
ps <- phyloseq(tax_table(tax), sample_data(samdf),
               otu_table(seqtab_final, taxa_are_rows = FALSE),
               phy_tree(fitGTR$tree))

if(nrow(seqtab_final) > nrow(sample_data(ps))){warning("Warning: All samples not included in phyloseq object, check sample names match the sample metadata")}

##save phyloseq object
saveRDS(ps, "output/rds/ps.rds")

#Output tables of results
dir.create("output/csv")
dir.create("output/otu_tables/unfiltered/")

##Export raw csv
speedyseq::psmelt(ps) %>%
  write.csv(file = "output/otu_tables/rawdata.csv")

seqateurs::summarise_taxa(ps, "species", "SampleID") %>%
  spread(key="SampleID", value="totalRA") %>%
  write.csv(file = "output/otu_tables/unfiltered/spp_sum.csv")

seqateurs::summarise_taxa(ps, "genus", "SampleID") %>%
  spread(key="SampleID", value="totalRA") %>%
  write.csv(file = "output/otu_tables/unfiltered/gen_sum.csv")
```


## Read in psyllid phylogeny
```{R psyllid phylogeny}
psyllid_tree <- read.tree(text=readLines("sample_data/psyllid_beast_tree.nwk"))

# Match names with alpha diversity
psyllid_tree$tip.label <- psyllid_tree$tip.label %>%
  str_squish() %>%
  str_replace_all(pattern="\\.", replacement=" ") %>%
  str_replace_all(pattern="Acizzia hakae", replacement="Acizzia hakeae") %>%
  str_replace_all(pattern="POLLENISLAND", replacement="POLLEN ISLAND") %>%
  str_replace_all(pattern="Ctenarytaina fuchsiae$", replacement="Ctenarytaina fuchsia A") %>%
  str_replace_all(pattern="Ctenarytaina fuchsiaeB", replacement="Ctenarytaina fuchsia B") %>%
  str_replace_all(pattern="Ctenarytaina fuchsiaeC", replacement="Ctenarytaina fuchsia C") %>%
  str_replace_all(pattern="Ctenarytaina clavata", replacement="Ctenarytaina clavata sp ") %>%
  str_replace_all(pattern="Ctenarytaina clavata sp $", replacement="Ctenarytaina clavata sp A") %>%
  str_replace_all(pattern="Ctenarytaina sp$", replacement="Ctenarytaina sp ") %>%
  str_replace_all(pattern="Ctenarytaina spA", replacement="Ctenarytaina sp A") %>%
    str_replace_all(pattern="Ctenarytaina spB", replacement="Ctenarytaina sp B") %>%
  str_replace_all(pattern="Ctenarytaina unknown", replacement="Ctenarytaina insularis") %>%  
  str_replace_all(pattern="Psylla apicalisA", replacement="Psylla frodobagginsi") %>%
  str_replace_all(pattern="Psylla apicalisB", replacement="Psylla apicalis") %>%
  str_replace_all(pattern="carmichaeliae", replacement="carmichaeliae ") %>%
  str_replace_all(pattern="Trioza sp", replacement="Trioza sp ") %>%
  str_replace_all(pattern="Trioza acutaB", replacement="Trioza acuta B") %>%
  str_replace_all(pattern="Trioza gourlay", replacement="Trioza gourlayi") %>%
  str_replace_all(pattern="BRENDAMAY", replacement="BRENDA MAY") %>%
  str_replace_all(pattern="PRICES", replacement="PRICES VALLEY") %>%  
  str_replace_all(pattern="Acizzia sp", replacement="Acizzia errabunda") %>% 
  str_replace_all(pattern=" ", replacement="_") %>%
  trimws(which="right")

# Subset to only those in sample data
psyllid_tree$tip.label[!psyllid_tree$tip.label %in% samdf$psyllid_spp]
pruned.tree <- drop.tip(psyllid_tree, psyllid_tree$tip.label[!psyllid_tree$tip.label %in% samdf$psyllid_spp] )
```

### Summarise taxonomic assignment

```{r sum taxa}
#Fraction of reads assigned to each taxonomic rank
sum_reads <- speedyseq::psmelt(ps) %>%
  gather("Rank","Name",rank_names(ps)) %>%
  group_by(Rank) %>% 
  mutate(Name = replace(Name, str_detect(Name, "__"),NA)) %>% # This line turns the "__" we added to lower ranks back to NA's
  dplyr::summarise(Reads_classified = sum(Abundance * !is.na(Name))) %>%
  mutate(Frac_reads = Reads_classified / sum(sample_sums(ps))) %>%
  mutate(Rank = factor(Rank, rank_names(ps))) %>%
  arrange(Rank)

#Fraction of ASV's assigned to each taxonomic rank
sum_otu <- tax_table(ps) %>%
  as("matrix") %>%
  as_tibble(rownames="OTU") %>%
  gather("Rank","Name",rank_names(ps)) %>%
  group_by(Rank) %>%
  mutate(Name = replace(Name, str_detect(Name, "__"),NA)) %>% # This line turns the "__" we added to lower ranks back to NA's
  dplyr::summarise(OTUs_classified = sum(!is.na(Name))) %>%
  mutate(Frac_OTUs = OTUs_classified / ntaxa(ps)) %>%
  mutate(Rank = factor(Rank, rank_names(ps))) %>%
  arrange(Rank)

print(sum_reads)
print(sum_otu)
```


## Curate with LULU

```{r LULU}
library(lulu)
# Make blast database from the ASV's
seqateurs::ps_to_fasta(ps, "output/all_taxa.fasta", seqnames = "sequence")
taxreturn::blast_install(dest.dir = "bin")

# Blast against themseleves
seqs <- insect::char2dna(colnames(seqtab_final))
names(seqs) <- colnames(seqtab_final)
matchlist <- blast(query=seqs, db="output/all_taxa.fasta", output_format = "tabular") %>%
  select(qseqid, sseqid, pident) %>%
  as.data.frame()

# Run lulu
otutab <- as.data.frame(t(seqtab_final))
curated_result <- lulu(otutab, matchlist)
saveRDS(curated_result, "lulu_curated.rds")

# get mapping
curated_result <- readRDS("lulu_curated.rds")
mapping <- curated_result[["otu_map"]] %>% 
  rownames_to_column("otu") %>%
  left_join(tax %>% as.data.frame %>% rownames_to_column("otu"), by="otu")

#assign_to ps
ps_curated <- ps
otu_table(ps_curated) <- otu_table(t(curated_result$curated_table), taxa_are_rows = FALSE)
ps_curated <- ps_curated %>% #Remove mocks
  filter_taxa(function(x) mean(x) > 0, TRUE) #Drop missing taxa from table 
message(ntaxa(ps) - ntaxa(ps_curated), " removed by LULU curation")

ps <- ps_curated


#Rename taxa
taxa_names(ps) <- paste0("SV", seq(ntaxa(ps)),"-",tax_table(ps)[,8])
```

## Remove outlier Samples

Detecting and potentially removing samples outliers (those samples with underlying data that do not conform to experimental or biological expectations) can be useful for minimizing technical variance. This can be caused by a number of reasons, including low-reads assigned to that sample. In this case we remove all samples below 1000 reads, as these include all samples contributing to lower than usual ASV counts.

Rarefaction curves are useful	to	assess	sensitivity	of	sample	size	to	observed	alpha-diversity estimates.
```{r replicate reproducibility}
## Remove mocks
rm_mocks <- c("mockA_S51", "MockEven_S193", "Mock_S192", "PCRctrl_S191", "MockStaggered_S194", "PCRctrl_S191", "91_S167")

#check mocks
ps0 <- ps %>% 
  subset_samples(!sample_names(ps) %in% rm_mocks) %>% #Remove mocks
  filter_taxa(function(x) mean(x) > 0, TRUE) #Drop missing taxa from table 

message((nsamples(ps) - nsamples(ps0)), " outlier samples dropped")

#Plot rarefaction curve
out <- rarecurve(otu_table(ps0), step=100)

rare <- lapply(out, function(x){
  b <- as.data.frame(x)
  b <- data.frame(OTU = b[,1], count = rownames(b))
  b$count <- as.numeric(gsub("N", "",  b$count))
  return(b)
})
names(rare) <- sample_names(ps1)

rare <- map_dfr(rare, function(x){
  z <- data.frame(x)
  return(z)
}, .id = "sample")

# threshold for read removal
threshold = 1000

gg.rare <- ggplot(data = rare)+
  geom_line(aes(x = count, y = OTU, group=sample), alpha=0.5)+
  geom_point(data = rare %>% group_by(sample) %>% top_n(1, count),aes(x = count, y = OTU, colour=(count > threshold))) +
  scale_x_continuous(labels =  scales::scientific_format()) +
  geom_vline(xintercept=threshold, linetype="dashed") +
  labs(colour = "> Read \n Threshold") +
  xlab("Sequence reads") +
  ylab("Observed ASV's")

gg.rare

#Remove all samples under the minimum read threshold 
ps1 <- prune_samples(sample_sums(ps0)>=threshold, ps0) 
ps1 <- filter_taxa(ps1, function(x) mean(x) > 0, TRUE) #Drop missing taxa from table
message(nsamples(ps) - nsamples(ps1), " Samples and ", ntaxa(ps) - ntaxa(ps1), " taxa under read threshold Dropped")
```

## Average dissimilarities

Calculate Average dissimilarities between PCR replicates, biological replicates of the same species, hostplants, and different species on same hostplants

```{r Avg Dissimiarity}
# Calculate overall distance
#Rarefy replicates to same read depth
nspec <- specnumber(otu_table(ps1)) # observed number of species
raremax <- min(rowSums(otu_table(ps1)))
rare <- rrarefy(otu_table(ps1), raremax)

dist <- vegan::vegdist(rare, method="bray") %>%
  as.matrix()

# Plot pca
r.pcx <- prcomp(dist)
factoextra::fviz_pca_biplot(r.pcx)

# Reformat to Long format
xy <- t(combn(colnames(dist), 2))
pw <- data.frame(xy, dist = dist[xy]) %>%
  set_colnames(c("SampleID.x", "SampleID.y", "dist")) %>%
  left_join(samdf %>% 
              dplyr::select(SampleID, Sample_Name, psyllid_spp, psyllid_genus, hostplant_spp, replicated) %>%
              dplyr::rename(SampleID.x = SampleID), by="SampleID.x")%>%
  left_join(samdf %>% 
              dplyr::select(SampleID, Sample_Name, psyllid_spp, psyllid_genus, hostplant_spp, replicated) %>%
              dplyr::rename(SampleID.y = SampleID), by="SampleID.y")


# Distance between replicates of same sample
repdist <- pw %>% filter(Sample_Name.x == Sample_Name.y)

# Distance between samples of same species
sppdist <- pw %>% filter(psyllid_spp.x == psyllid_spp.y)

# Distance between psyllids on same hostplat
#      psyllopsis fraxini 94- 107 - 113
# psyllopsis fraxinicola 93 - 106 - 112
# 93+94 were the same tree, 106+107 on the same tree and 112+113 were the same treee
hostplantdist <- pw %>% filter(psyllid_spp.x == psyllid_spp.y)

dist <- ps1 %>% 
    subset_samples(Sample_Name %in%  c("93", "94","106", "107","112", "113")) %>%
    filter_taxa(function(x) mean(x) > 0, TRUE) %>%
    otu_table()  %>%
    vegan::vegdist(method="jaccard")

factoextra::fviz_pca(prcomp(dist))

# Otther set
#psylla apicalis 200 - 201
#psylla frodobagginsi 200 - 201 #in this case they are named after their tree - ie 200 and 200 were same tree, i shoul drename thes ein some way

dist2 <- ps1 %>% 
    subset_samples(Sample_Name %in%  c("200", "94","106", "107","112", "113")) %>%
    filter_taxa(function(x) mean(x) > 0, TRUE) %>%
    otu_table()  %>%
    vegan::vegdist(method="jaccard")

# could i do some cluster analysis here?

# Or is adonis besst
#adonis(~ psyllid_spp, method = "euclidean", data = metadata))
```

### Merge technical replicates

With only 16 samples replicated, and only on the first 2 sequencing runs i dont think there is a way to explicitly take into account replicate variability, therefore all replicates were merged

```{r replicates}
# Merge replicates
ps.merged <- ps1 %>%
    merge_samples(group = "Sample_Name")

#This loses the sample metadata - Need to add it agian
sample_data(ps.merged) <- sample_data(ps1) %>%
  filter(!duplicated(Sample_Name)) %>%
  magrittr::set_rownames(.$Sample_Name)

#Drop missing taxa from table
ps.merged <- filter_taxa(ps.merged, function(x) mean(x) > 0, TRUE)
```

## Prevalence assesment

View prevalence of different phyla
```{r prevalence-assessment}
# Calculate taxon prevalence across the data set
prevdf <- apply(X = otu_table(ps.merged), MARGIN = ifelse(taxa_are_rows(ps.merged), yes = 1, no = 2), FUN = function(x){sum(x > 0)})

# Add taxonomy and total read counts to prevdf - change this to tidyverse code
prevdf <- data.frame(Prevalence = prevdf, TotalAbundance = taxa_sums(ps.merged), tax_table(ps.merged))

#Prevalence plot
gg.prev <- subset(prevdf, phylum %in% get_taxa_unique(ps0, "phylum")) %>%
  ggplot(aes(TotalAbundance, Prevalence / nsamples(ps.merged),color=family)) +
  geom_point(size = 3, alpha = 0.7) +
  scale_x_log10() +
  xlab("Total Abundance") + ylab("Prevalence [Frac. Samples]") +
  facet_wrap(~phylum) +
  theme(legend.position="none") +
  ggtitle("Phylum Prevalence in All Samples\nColored by Family")

gg.prev
```

## Taxon Filtering

Remove all non-bacterial taxa, mitochondria, chloroplast and cyanobacteria.
We also remove all taxa contained within the blank sample from other samples? - Check these

```{r taxon-cleaning}
get_taxa_unique(ps.merged, "domain")

ps.merged # Check the number of taxa prior to removal
ps2 <- ps.merged %>%
  subset_taxa(
    domain == "Bacteria" & #This is probably the only required one
    family  != "Mitochondria" &
    order   != "Chloroplast" &
    phylum != "Cyanobacteria"
  )
ps2 # Confirm that the taxa were removed
get_taxa_unique(ps2, "phylum")
get_taxa_unique(ps2, "class")
get_taxa_unique(ps2, "family")
```

# Alpha diversity 


## Summary of most common taxonomic groups

 Keep shannons, raw asv's and PD. 
 
Tests of community clusterign could be moved later - see: https://www.nature.com/articles/s41467-019-10191-3

alpha diversity (observed, phylogenetic,  alpha-diversity metrics like Shannon/SImpson etc)

In order to further test this notion and to assess how environmental filtering may differ among host clades, we utilized two ecophylogenetics analyses: mean phylogenetic distance (MPD) and mean nearest taxon distance (MNTD). These tests assess the degree of phylogenetic clustering within each sample (host) relative to a permuted null model. Assuming phylogenetic niche conservatism (i.e., closely related taxa overlap along niche axes), then host diet or gut physiology may select for phylogenetically clustered taxa with overlapping niches, while in the absence of such strong selection, competition via niche conservatism would lead to phylogenetic overdispersion40. Phylogenetic overdispersion may also result from facilitation (i.e., beneficial microbe–microbe interactions), such as when distantly related taxa form consortia to break down complex plant polymers40. MPD is more sensitive to overall patterns of phylogenetic clustering and evenness, while MNTD is more sensitive to patterns at the tree tips31.

## Alpha diversity metrics

Test this with rarefied samples to see the robustness to sequencing depth variation
```{r alpha diversity}
phy <- phyloseq::phy_tree(ps2)
phydist <- cophenetic(phy)
comm <- as(phyloseq::otu_table(ps2), "matrix") 

# Get richness measures
richness <- phyloseq::estimate_richness(ps2, measures=c("Chao1", "Shannon", "Simpson")) %>%
  rownames_to_column("Sample_Name") %>%
  mutate(Sample_Name = Sample_Name %>% 
           str_remove("^X") %>%
           str_replace_all("\\.", " "))

#Set number of randomisations for calculating significance
randoms <- 10

# Calculate Faith's PD-index & Species richness - with Standard errors
sespd <- picante::ses.pd(comm, phy, null.model = "taxa.labels", include.root = F, runs = randoms)

# Weighted mpd
w_mpd <- picante::ses.mpd(comm, phydist, null.model = "taxa.labels", abundance.weighted = TRUE, runs = randoms)

#Unweighted mpd
u_mpd <- picante::ses.mpd(comm, phydist, null.model = "taxa.labels", abundance.weighted = FALSE, runs = randoms) 

# Weighted mntd
w_mntd <- picante::ses.mntd(comm, phydist, null.model = "taxa.labels", abundance.weighted = TRUE, runs = randoms) 

# Unweighted NRI
u_mntd <- picante::ses.mntd(comm, phydist, null.model = "taxa.labels", abundance.weighted = FALSE, runs = randoms) 

# Join together
div_table <- sespd %>%
  rownames_to_column("Sample_Name") %>%
  dplyr::select(Sample_Name, ntaxa, pd.obs.z) %>%
  bind_cols(w_mpd %>% dplyr::select(contains("obs.z"))) %>%
  bind_cols(u_mpd %>% dplyr::select(contains("obs.z"))) %>%
  bind_cols(w_mntd %>% dplyr::select(contains("obs.z"))) %>%
  bind_cols(u_mntd %>% dplyr::select(contains("obs.z"))) %>%
  dplyr::rename(alpha = ntaxa, pd = pd.obs.z,
         w_mpd = mpd.obs.z, u_mpd = mpd.obs.z1,
         w_mntd = mntd.obs.z, u_mntd = mntd.obs.z1)  %>%
  left_join(richness, by="Sample_Name") %>%
  left_join(sample_data(ps2) %>% 
              as.data.frame() %>%
              filter(!duplicated(Sample_Name)) %>%
              dplyr::select(Sample_Name, psyllid_spp),
            by = "Sample_Name") 
# Summarise means
div_table %>%
  summarise_if(is.numeric, mean)
```


## Association with phylogeny 

Here we calculate the phylogenetic signal in the alpha diversity and community assembly metrics, or the extent to which these values are statistically related to phylogeny.

see supplementary 15 - https://static-content.springer.com/esm/art%3A10.1038%2Fs41467-019-10191-3/MediaObjects/41467_2019_10191_MOESM1_ESM.pdf


## Calculate phylogenetic signal

```{r Phylosignal}
dat <- div_table  %>%
  filter(psyllid_spp %in% pruned.tree$tip.label) %>% #Subset to common 
  group_by(psyllid_spp) %>%
  dplyr::select(-Sample_Name) %>%
  summarise_all(mean) %>%
  arrange(match(psyllid_spp, pruned.tree$tip.label)) %>%
  as.data.frame() %>%
  magrittr::set_rownames(.$psyllid_spp) %>%
  dplyr::select(-psyllid_spp)

# Add positive and negative controls
dat$random <- rnorm(length(dat$alpha), sd = 10) #Random association
dat$bm <- rTraitCont(pruned.tree) #Brownian motion

# Make phylosignal object and measure signal between univariate traits.
p4d <- phylobase::phylo4d(pruned.tree, dat)	
signal <- phylosignal::phyloSignal(p4d = p4d, method = "all")	

# Plots
barplot.phylo4d(p4d, tree.type = "phylo", tree.ladderize = TRUE)

# Locate signal
lipa <- lipaMoran(p4d)
lipa.p4d <- lipaMoran(p4d, as.p4d = TRUE)

barplot.phylo4d(p4d, bar.col=(lipa$p.value < 0.05) + 1, center = FALSE , scale = FALSE)

barplot.phylo4d(lipa.p4d, bar.col = (lipa$p.value < 0.05) + 1, center = FALSE, scale = FALSE)
```	


## Phylosymbiosis

### Prepare distance matrices
```{r prepare distances}
## Psyllid phylogeny cophenetic distance matrix
phylo.dist <- cophenetic(pruned.tree) %>%
  as.data.frame() %>%
  rownames_to_column("psyllid_spp.x") %>%
  pivot_longer(cols=-psyllid_spp.x,
               names_to="psyllid_spp.y",
               values_to = "dist") %>%
  right_join(sample_data(ps2) %>%
              as_tibble() %>%
              dplyr::select(Sample_Name, psyllid_spp) %>%
              dplyr::rename(psyllid_spp.x = psyllid_spp, Sample_Name.x = Sample_Name),
            by="psyllid_spp.x")%>%
  right_join(sample_data(ps2) %>%
              as_tibble() %>%
              dplyr::select(Sample_Name, psyllid_spp) %>%
              dplyr::rename(psyllid_spp.y = psyllid_spp, Sample_Name.y = Sample_Name),
            by="psyllid_spp.y") %>%
  filter(!is.na(dist))%>%
  dplyr::select(-psyllid_spp.y, -psyllid_spp.x) %>%
  pivot_wider(names_from = Sample_Name.y, values_from = dist) %>%
  magrittr::set_rownames(.$Sample_Name.x) %>%
  dplyr::select(-Sample_Name.x)%>%
  as.matrix()

# COI Genetic distance matrix
coi.dist <- read_csv("sample_data/genetic_distance.csv", col_names = TRUE) %>%
  column_to_rownames("SampleID") 
coi.dist[upper.tri(coi.dist)] <- 0
coi.dist<- coi.dist + t(coi.dist)
coi.dist[is.na(coi.dist)] <- 0

coi.dist <- coi.dist %>%
  magrittr::set_colnames(rownames(.) %>% str_replace(pattern="\\_S(.*)$",replacement="") %>% make.unique()) %>%
  rownames_to_column("SampleID") %>%
  mutate(SampleID = SampleID %>% str_replace(pattern="\\_S(.*)$",replacement="") %>% make.unique()) %>%
  column_to_rownames("SampleID") %>%
  as.matrix()

# Hostplant genetic distance matrix
plant.dist <- read_csv("sample_data/plant_distance.csv") %>%
  dplyr::select(-one_of("X1", "Host plants")) %>%
  column_to_rownames("SampleID") %>%
  magrittr::set_colnames(rownames(.) %>% str_replace(pattern="\\_S(.*)$",replacement="") %>% make.unique()) %>%
  rownames_to_column("SampleID") %>%
  mutate(SampleID = SampleID %>% str_replace(pattern="\\_S(.*)$",replacement="") %>% make.unique()) %>%
  column_to_rownames("SampleID") %>%
  as.matrix()

# Spatial distance matrix
envData <- sample_data(ps2) %>%
  as_data_frame() %>%
  dplyr::select(Sample_Name, lat, long) %>%
  tibble::column_to_rownames("Sample_Name") %>%
  drop_na()

spat.dist <- spDists(as.matrix(envData), longlat=TRUE) %>%
  as.data.frame() %>%
  magrittr::set_rownames(rownames(envData) %>% str_replace(pattern="\\_S(.*)$",replacement="") %>% make.unique()) %>%
  magrittr::set_colnames(rownames(envData) %>% str_replace(pattern="\\_S(.*)$",replacement="") %>% make.unique()) %>%
  rownames_to_column("SampleID") %>%
  mutate(SampleID = SampleID %>% str_replace(pattern="\\_S(.*)$",replacement="") %>% make.unique()) %>%
  column_to_rownames("SampleID") %>%
  as.matrix()
```

## Whole dataset

The dendrogram approach first produces a hierarchical dendrogram depicting microbiota beta diversity using the UPGMA. It then quantifies the match between the microbial dendrogram and the host phylogeny using the normalized Robinson-Foulds metric. Significance is based on 999 permutations of host tip label phylogeny. Note that this is different from the original dendrogram-based approach that uses random phylogenies with the same numbers of tips as null expectations. However, the two null models gave nearly exactly the same results (Fig. S8), so we present only the tip-shuffling null model in the main text. T
see: https://github.com/FloMazel/Phylosymbiosis-Ecological-model

http://www.phytools.org/static.help/multiRF.html

But also see how this paper has done it with PACO https://www.nature.com/articles/s41467-019-10191-3.pdf Using random subsampling? Then pull out the most associated also file:///C:/Users/ap0y/Dropbox/workprojects/Papers/10.1002_ecy.1955.pdf

```{r phylosymbiosis}
dir.create("output/phylosymbiosis")
# Get OTU tables
otutab <- otu_table(ps2)
#Impute zeroes for compositional distances
otutab_n0 <- as.matrix(zCompositions::cmultRepl(otutab, method="CZM", output="prop"))
#Root phylogenetic tree
phy_tree(ps2) <- multi2di(phy_tree(ps2))

#Calculate different distance metrics
metrics <- c("Bray", "Jaccard", "Aitchison","Philr", "Unifrac", "WUnifrac") 
distlist <- vector("list", length=length(metrics))
names(distlist) <- metrics

distlist$Jaccard <- as.matrix(vegdist(otutab, method="jac",binary = T))
distlist$Bray <- as.matrix(vegdist(otutab, method="bray"))
distlist$Aitchison <- as.matrix(vegdist(CoDaSeq::codaSeq.clr(otutab_n0), method="euclidean"))
distlist$Philr <- as.matrix(vegdist(philr::philr(otutab_n0, phy_tree(ps2),
                                                part.weights='enorm.x.gm.counts',
                                                ilr.weights='blw.sqrt'), method="euclidean"))
distlist$Unifrac <- as.matrix(phyloseq::UniFrac(ps2, weighted=FALSE, parallel = TRUE))
distlist$WUnifrac <- as.matrix(phyloseq::UniFrac(ps2, weighted=TRUE, parallel = TRUE))

# Adonis test
metadata <- sample_data(ps2) %>%
  as_data_frame()
adonis_results <- distlist %>%
  purrr::map(function(x) {
    bind_rows(
    broom::tidy(adonis(x~psyllid_spp, method="euclidean", data=metadata)$aov.tab) %>% dplyr::slice(1),
    broom::tidy(adonis(x~hostplant_spp, method="euclidean", data=metadata)$aov.tab) %>% dplyr::slice(1)
    )
})  %>%
  bind_rows(.id="dist")

write_csv(adonis_results, "output/phylosymbiosis/adonis_fulldata.csv")

# Matrix correlations
#only use samples present in all
samples <- Reduce(intersect, list(rownames(otu_table(ps2)), colnames(phylo.dist), colnames(coi.dist), colnames(plant.dist), colnames(spat.dist)))

# Mantel test
mantel_results <- distlist %>%
  purrr::map(function(x){
    run_mantel(x, dists=c("phylo.dist","coi.dist","plant.dist", "spat.dist"),
               samples=samples, type="mantel")
  }) %>%
  bind_rows(.id="dist") 

write_csv(mantel_results, "output/phylosymbiosis/mantel_fulldata.csv")

# Partial Mantel Test
pmantel_results <- distlist %>%
  purrr::map(function(x){
    run_mantel(x, dists=c("phylo.dist","coi.dist","plant.dist", "spat.dist"),
               samples=samples, type="partial")
  }) %>%
  bind_rows(.id="dist") 

write_csv(pmantel_results, "output/phylosymbiosis/pmantel_fulldata.csv")

# Topological comparison metrics
topo_results <- distlist %>%
  purrr::map(function(x){
    run_tree_comparison(x, dists=c("phylo.dist","coi.dist","plant.dist", "spat.dist"), #Change these to matrices
               samples=samples,
               clust.method="average",
               measure="all",
               runs=100)
  }) %>%
  bind_rows(.id="metric") 
write_csv(topo_results, "output/phylosymbiosis/topo_fulldata.csv")

## Visualise what these look like
#tree1  <- as.phylo(hclust(as.dist(phylo.dist[samples,samples]), method = "ward.D2"))
#tree2 <- as.phylo(hclust(as.dist(distlist$Aitchison[samples,samples]), method = "ward.D2"))
#VisualizeMatching(RobinsonFoulds,tree1, tree2)

```

### Without Gammaproteobacteria
```{r phylosymbiosis no entero}
## TRY WITHOUT Gammaproteobacteria
ps2_subset <- ps2 %>%
 subset_taxa(class != "Gammaproteobacteria") %>% #is this working?
 filter_taxa(function(x) mean(x) > 0, TRUE)#Drop missing taxa from table
ps2_subset <- prune_samples(sample_sums(ps2_subset) >0 , ps2_subset)
message(nsamples(ps2_subset) - nsamples(ps2_subset), " Samples and ", ntaxa(ps2) - ntaxa(ps2_subset), " taxa under read threshold Dropped")

# Get OTU tables
otutab <- otu_table(ps2_subset)
#Impute zeroes for compositional distances
otutab_n0 <- as.matrix(zCompositions::cmultRepl(otutab, method="CZM", output="prop", suppress.print=quiet))
#Root phylogenetic tree
phy_tree(ps2_subset) <- multi2di(phy_tree(ps2_subset))

#Calculate different distance metrics
metrics <- c("Bray", "Jaccard", "Aitchison", "Unifrac", "WUnifrac") # "Philr"
distlist <- vector("list", length=length(metrics))
names(distlist) <- metrics

#Calculate different distance metrics
metrics <- c("Bray", "Jaccard", "Aitchison","Philr", "Unifrac", "WUnifrac") 
distlist <- vector("list", length=length(metrics))
names(distlist) <- metrics

distlist$Jaccard <- as.matrix(vegdist(otutab, method="jac",binary = T))
distlist$Bray <- as.matrix(vegdist(otutab, method="bray"))
distlist$Aitchison <- as.matrix(vegdist(CoDaSeq::codaSeq.clr(otutab_n0), method="euclidean"))
distlist$Philr <- as.matrix(vegdist(philr::philr(otutab_n0, phy_tree(ps2),
                                                part.weights='enorm.x.gm.counts',
                                                ilr.weights='blw.sqrt'), method="euclidean"))
distlist$Unifrac <- as.matrix(phyloseq::UniFrac(ps2, weighted=FALSE, parallel = TRUE))
distlist$WUnifrac <- as.matrix(phyloseq::UniFrac(ps2, weighted=TRUE, parallel = TRUE))

# Adonis test
metadata <- sample_data(ps2_subset) %>%
  as_data_frame()
adonis_results <- distlist %>%
  purrr::map(function(x) {
    bind_rows(
    broom::tidy(adonis(x~psyllid_spp, method="euclidean", data=metadata)$aov.tab) %>% dplyr::slice(1),
    broom::tidy(adonis(x~hostplant_spp, method="euclidean", data=metadata)$aov.tab)%>% dplyr::slice(1)
    )
})  %>%
  bind_rows(.id="dist")

write_csv(adonis_results, "output/phylosymbiosis/adonis_subset.csv")

# Matrix correlations
#only use samples present in all
samples <- Reduce(intersect, list(rownames(otu_table(ps2_subset)), colnames(phylo.dist), colnames(coi.dist), colnames(plant.dist), colnames(spat.dist)))

# Mantel test
mantel_results <- distlist %>%
  purrr::map(function(x){
    run_mantel(x, dists=c("phylo.dist","coi.dist","plant.dist", "spat.dist"),
               samples=samples, type="mantel")
  }) %>%
  bind_rows(.id="dist") 
write_csv(mantel_results, "output/phylosymbiosis/mantel_subset.csv")

# Partial Mantel Test
pmantel_results <- distlist %>%
  map(function(x){
    run_mantel(x, dists=c("phylo.dist","coi.dist","plant.dist", "spat.dist"),
               samples=samples, type="partial")
  }) %>%
  bind_rows(.id="dist") 
write_csv(pmantel_results, "output/phylosymbiosis/pmantel_subset.csv")

# Topological comparison metrics
topo_results <- distlist %>%
  purrr::map(function(x){
    run_tree_comparison(x, dists=c("phylo.dist","coi.dist","plant.dist", "spat.dist"), #Change these to matrices
               samples=samples,
               clust.method="average",
               measure="all",
               runs=100)
  }) %>%
  bind_rows(.id="metric") 
write_csv(topo_results, "output/phylosymbiosis/topo_subset.csv")

```

### BDTT

```{r timeslice}
library(TreeDist)
#Look at depth categories in tree
hist(get_all_node_depths(phy_tree(ps2)))

#Set slices
slices=seq(0,1.6,0.1)

treelist <- vector("list", length=length(slices))
for (i in 1:length(slices)){
  newps <- speedyseq::tree_glom(ps2, resolution = slices[i])
  treelist[[i]] <- phy_tree(newps)
}
class(treelist) <- "multiPhylo"
names(treelist) <- paste0("Timeslice = ",slices)
gg.tree <- ggtree(treelist) + facet_wrap(~.id, scale="free", nrow=1) + theme_tree2()

Betas <- BDTT(ps2, slices, metrics=c("Bray", "Aitchison", "Philr"), zeroes="Impute", cores=1, quiet=FALSE) #"Unifrac", "WUnifrac"

# Matrix correlations
#only use samples present in all
samples <- rownames(otu_table(ps2))
samples <- samples[(samples %in% colnames(phylo.dist)) &
                              (samples %in% colnames(coi.dist))  &
                              (samples %in% colnames(plant.dist)) &
                              (samples %in% colnames(spat.dist))]

# MANTEL
BTDD_mantel <- Betas %>%
  purrr::map_depth(.depth=2, function(x){
    run_mantel(x, dists=c("phylo.dist","coi.dist","plant.dist", "spat.dist"),
               samples=samples, type="mantel")
  }) %>%
  purrr::map(bind_rows,.id="metric") %>%
  bind_rows(.id="slice")


gg.mantel <- BTDD_mantel %>%
  mutate(Comparison = paste0("microbe~", dist1)) %>%
  ggplot(aes(y=statistic, x=slice, colour=Comparison, group=factor(Comparison)))+ 
  geom_line(linetype="dashed") +
  geom_line(data=BTDD_mantel %>% mutate(Comparison = paste0("microbe~",dist1)) %>% mutate(statistic = case_when (signif > 0.01 ~ as.double(NA), signif < 0.01 ~ statistic)), linetype="solid") +
  geom_hline(yintercept=0) +
  facet_wrap(~metric) +
     #theme_classic2()   +
     theme(legend.position = "right",
           panel.grid.major.y = element_line(colour="grey", size=0.5, linetype="dashed"),
           strip.background = element_rect(fill = "grey92", 
                                           colour = "black", size = 1)
     )  + scale_color_brewer(palette="Paired")

gg.mantel 

# Partial Mantel
BTDD_pm <- Betas %>%
  purrr::map_depth(.depth=2, function(x){
    run_mantel(x, dists=c("phylo.dist","coi.dist","plant.dist", "spat.dist"),
               samples=samples, type="partial")
  }) %>%
  purrr::map(bind_rows,.id="dist") %>%
  bind_rows(.id="metric")

gg.pmantel <- BTDD_pm %>%
  mutate(Comparison = paste0("microbe~",dist1,"~",dist2)) %>%
  ggplot(aes(y=statistic, x=slice, colour=Comparison, group=factor(Comparison)))+ 
  geom_line(linetype="dashed") +
  geom_line(data=BTDD_pm %>% mutate(Comparison = paste0("microbe~",dist1,"~",dist2)) %>% mutate(statistic = case_when (signif > 0.01 ~ as.double(NA), signif < 0.01 ~ statistic)), linetype="solid") +
  geom_hline(yintercept=0) +
  facet_wrap(~metric) +
     #theme_classic2()   +
     theme(legend.position = "right",
           panel.grid.major.y = element_line(colour="grey", size=0.5, linetype="dashed"),
           strip.background = element_rect(fill = "grey92", 
                                           colour = "black", size = 1)
     )  + scale_color_brewer(palette="Paired")

gg.pmantel

# Topological comparison
BTDD_topo <- Betas %>%
  purrr::map_depth(.depth=2, function(x){
    run_tree_comparison(x, dists=c("phylo.dist","coi.dist","plant.dist", "spat.dist"), #Change these to matrices
               samples=samples,
               clust.method="average",
               measure="all",
               runs=100)
  }) %>%
  purrr::map(bind_rows,.id="dist") %>%
  bind_rows(.id="metric")

write_csv(BTDD_topo, "output/phylosymbiosis/BTDD_topo.csv")

# Plot topo
```

# Cophylogeny
Due to the association with close branches we looked for patterns of cophylogeny

ggtree tanglegrams: https://stackoverflow.com/questions/57161083/coloring-lines-in-tanglegram-based-on-position-of-nodes

## Whole dataset - PACO

```{r PACO}
library(paco)

# Merge species for beta diversity
ps.sppmerged <- ps2 %>%
    merge_samples(group = "psyllid_spp", fun=mean)

#This loses the sample metadata - Need to add it agian
sample_data(ps.sppmerged) <- sample_data(ps2) %>%
  filter(!duplicated(psyllid_spp)) %>%
  magrittr::set_rownames(.$psyllid_spp)

ps3 <- filter_taxa(ps.sppmerged, function(x) mean(x) > 0, TRUE) #Drop missing taxa from table

#Prepare co-occurance matrix
otu <- ps3 %>%
    subset_samples(sample_names(ps3) %in% pruned.tree$tip.label) %>%
    otu_table %>%
    as.matrix() %>% 
  apply(2, function(x) ifelse(x > 0, 1, 0))

# Get cophenetic distance
h_dist <- cophenetic(pruned.tree)
s_dist <- cophenetic(phy_tree(ps3))

# prepare paco data
D <- prepare_paco_data(H=h_dist, P=s_dist, HP=otu)
# Add pcord
D <- add_pcoord(D, correction='cailliez')

D$H_PCo %>% as.vector %>% summary %>% print
D$P_PCo %>% as.vector %>% summary %>% print

# plot Principal coordinates
D$H_PCo <-  neg_to_pos_eig(D$H_PCo)
D$P_PCo <-  neg_to_pos_eig(D$P_PCo)

tmp <- D$H_PCo[,c('Axis.1', 'Axis.2')] %>% as.data.frame
p_host <- ggplot(tmp, aes(Axis.1, Axis.2)) +
    geom_point() +
    theme_bw() +
  ggtitle("Host PCA")
#One really to the left is Atmetocranium.myersi

tmp <- D$P_PCo[,c('Axis.1', 'Axis.2')] %>% as.data.frame
p_micro <- ggplot(tmp, aes(Axis.1, Axis.2))  +
    geom_point() +
    theme_bw()+
  ggtitle("Microbiome PCA")
    
plot(p_host + p_micro)

# run paco
paco_run <- PACo(D, nperm=9, seed=909, method='quasiswap', symmetric=TRUE) #Symetric - is one meant to track the evolution of another?

#print overall significance
print(paco_run$gof)

# Get interaction-specific cophylogenetic contributions using leave-one-out jacknifing
# whereby one link is removed and the fit reassessed.
#The contribution of that link is therefore the change in fit whensaid link is removed.
paco_run <- paco_links(paco_run, .parallel = TRUE)

#Get observed residuals of Procrustean superimposition 
#small residual indicates a link more congruent with a cophylogenetic hypothesis
paco_residuals <- residuals_paco(paco_run$proc, type = "interaction")

# Visualise residuals
res <- data.frame(OTU=names(paco_residuals), values=unname(paco_residuals)) %>%
  separate(OTU, into=c("Sample_Name", "OTU"), sep="-", extra="merge") %>%
  left_join(as.data.frame(tax_table(ps3)) %>%
              rownames_to_column("OTU"), by="OTU") 

ggplot(res, aes(x=values))+
  geom_density(fill='grey70')+
  facet_wrap(~phylum) +
  theme_bw()+
  xlab('Procrustes residuals')+
  ylab('Frequency')


#Model bith the residuals and links by something, phylum?
simple_model <- lm(values ~ phylum, data=res)
summary(simple_model)

```

Could you then use a regression to look at the influence on results by bacterial taxonomy? ie Family?

## Carsonella cophylogeny

Definitely need to do the top carsonella, but need some flexibility in keeping the related ones, so top - 20% abundance? 

```{r subsetting}
#Flag top abundance carsonella epr sample
top_carson <- ps2 %>%
  transform_sample_counts(function (x) x/sum(x)) %>%
  speedyseq::psmelt() %>%
  filter(genus=="Candidatus Carsonella") %>%
  group_by(Sample) %>%  
  filter(Abundance > 0) %>%
  top_n(1, wt=Abundance) %>%
  mutate(top = TRUE) 

## Flag switched only if that OTU is present somewhere else on the seqrun and isnt top 
switched_carson <- carson %>%
    filter(Abundance > 0) %>%
    dplyr::select(OTU, top, seqrun, Abundance) %>%
    unique() %>%
    group_by(OTU, seqrun) %>%
    add_tally() %>%
    mutate(switched = case_when(
      n>1 & top== FALSE  ~ TRUE,
      n>1 & top== TRUE ~ FALSE,
      n==1 ~ FALSE,
    ))

otu <- ps2 %>%
  transform_sample_counts(function (x) x/sum(x)) %>%
  speedyseq::psmelt() %>%
  #filter(genus=="Candidatus Carsonella") %>%
  left_join(top_carson) %>%
  filter(genus=="Candidatus Carsonella", top==TRUE) %>%
  dplyr::select(OTU, psyllid_spp, SampleID, Abundance) %>%
  dplyr::group_by(OTU, psyllid_spp) %>%
    summarise(Abundance = sum(Abundance)) %>%
  pivot_wider(id_cols = psyllid_spp,
              names_from = OTU,
              values_from=Abundance,
              values_fill = list(Abundance = 0))  %>%
  column_to_rownames("psyllid_spp") %>%
    as.matrix() %>% 
  apply(2, function(x) ifelse(x > 0, 1, 0))

# Get cophenetic distance
h_dist <- cophenetic(pruned.tree)
s_dist <- ps3 %>%
    subset_taxa(rownames(tax_table(ps3)) %in% colnames(otu)) %>%
    phy_tree() %>%
  cophenetic

#Prepare co-occurance matrix
samples <- intersect(rownames(otu), rownames(h_dist))

otu <- subset(otu, rownames(otu) %in% samples)
# prepare paco data
D <- prepare_paco_data(H=h_dist[samples, samples], P=s_dist, HP=otu)
# Add pcord
D <- add_pcoord(D, correction='cailliez')

D$H_PCo %>% as.vector %>% summary %>% print
D$P_PCo %>% as.vector %>% summary %>% print

D$H_PCo <-  neg_to_pos_eig(D$H_PCo)
D$P_PCo <-  neg_to_pos_eig(D$P_PCo)

tmp <- D$H_PCo[,c('Axis.1', 'Axis.2')] %>% as.data.frame
p_host <- ggplot(tmp, aes(Axis.1, Axis.2)) +
    geom_point() +
    theme_bw() +
  ggtitle("Host PCA")

tmp <- D$P_PCo[,c('Axis.1', 'Axis.2')] %>% as.data.frame
p_micro <- ggplot(tmp, aes(Axis.1, Axis.2))  +
    geom_point() +
    theme_bw()+
  ggtitle("Microbiome PCA")
    
plot(p_host + p_micro)

# run paco
paco_run <- PACo(D, nperm=9, seed=909, method='quasiswap', symmetric=TRUE) #Symetric - is one meant to track the evolution of another?

#print overall significance
print(paco_run$gof)

# Get interaction-specific cophylogenetic contributions using jacknifing
# whereby one link is removed and the fit reassessed.
#The contribution of that link is therefore the change in fit whensaid link is removed.
paco_run <- paco_links(paco_run)

#Get observed residuals of Procrustean superimposition 
#small residual indicates a link more congruent with a cophylogenetic hypothesis
paco_residuals <- residuals_paco(paco_run$proc, type = "interaction")

# Visualise residuals
res <- data.frame(OTU=names(paco_residuals), values=unname(paco_residuals)) %>%
  separate(OTU, into=c("Sample_Name", "OTU"), sep="-", extra="merge") %>%
  left_join(as.data.frame(tax_table(ps3.carson)) %>%
              rownames_to_column("OTU"), by="OTU")
ggplot(res, aes(x=values))+
  geom_density(fill='grey70')+
  theme_bw()+
  xlab('Procrustes residuals')+
  ylab('Frequency')

otu.lut <- which(otu ==1, arr.ind=TRUE)
assoc <- cbind(rownames(otu)[otu.lut[,1]], colnames(otu)[otu.lut[,2]])

# Cant have - or space characters
assoc[,1] <- str_replace_all(assoc[,1], " ", "_")
assoc[,2] <- str_replace_all(assoc[,2], "-", "_")
assoc[,2] <- str_replace_all(assoc[,2], " ", "_")

newtree <- pruned.tree
newtree$tip.label <- str_replace_all(newtree$tip.label, " ", "_")

new_stree <- phy_tree(ps3.carson)
new_stree$tip.label <- new_stree$tip.label %>%
   str_replace_all( "-", "_") %>%
   str_replace_all( " ", "_")

library(phytools)
obj <- cophylo(tr1=newtree, tr2=new_stree, assoc=assoc, rotate=TRUE) 
plot.cophylo(obj, link.lwd=1, link.lty=2, fsize=0.5, pts=FALSE, link.type="curved")
dev.off() #reset graphics device

# The reason some of these are off are due to mismatches between the tree and the sample data
```

## Psyllid hostplant cophylogeny

Tanglegram of all plants and  
```{r}
library(rotl)
plants <- unique(plant_distance$`Host plants`)
resolved_names <- tnrs_match_names(plants[1])

```

# Figures

## Barplot

Use the means and standard errors? of the beta diversities by sample

Merge species, arrange microbiome by phylogeny, try the mantel tests on the merged data as well (but the seperate is probs more statistically better)

RESULTS: while there were significant differences between psyllid species there were no significant associations with phylogeny
```{r Figure 1}
# Merge species for beta diversity
ps.sppmerged <- ps2 %>%
    merge_samples(group = "psyllid_spp", fun=mean)

#This loses the sample metadata - Need to add it agian
sample_data(ps.sppmerged) <- sample_data(ps2) %>%
  filter(!duplicated(psyllid_spp)) %>%
  column_to_rownames("psyllid_spp")

ps3 <- filter_taxa(ps.sppmerged, function(x) mean(x) > 0, TRUE) #Drop missing taxa from table

#Rename taxa
taxa_names(ps3) <- paste0("SV", seq(ntaxa(ps3)),"-",tax_table(ps3)[,6])

ps3_bar <- #subset_taxa(ps3, !Family == "Enterobacteriaceae") %>%
            ps3 %>%
  speedyseq::tax_glom(taxrank = "Order") %>%           # agglomerate at Order level
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  speedyseq::psmelt() %>%
  mutate(plotlabel = Phylum) %>%
  mutate(plotlabel = case_when(
    Abundance >= 0.01 & Phylum=="Proteobacteria" ~ paste0(Phylum, " - ", Order), # Change this to whatever taxrank we want
    Abundance >= 0.01 & !Phylum=="Proteobacteria"~ Phylum ,
    Abundance < 0.01 ~ "NA"
    )) %>%
  dplyr::na_if("NA") %>%
  select(Sample, plotlabel, Phylum, Order, Abundance)

# Plot them together
library(ggtree)
library(treeio)
p <- ggtree(pruned.tree) + geom_tiplab(align=TRUE) + geom_nodelab(geom='label') +
    scale_x_continuous(expand=c(0, 0.1)) + theme_tree2()


Fig1 <- facet_plot(p, panel='Relative Abundance', data = ps3_bar %>% mutate(id = Sample), geom=ggstance::geom_barh,
                 mapping=aes(x = Abundance, fill = plotlabel), stat="identity")  + 
  scale_fill_manual(values=colorRampPalette(brewer.pal(11, "Set1"))(length(unique(ps3_bar$plotlabel))-1), na.value="grey")


Fig1 <- facet_plot(Fig1, panel='Species Richness', data = bt_randomdf %>% mutate(id = psyllid_spp),
                          geom=ggstance::geom_pointrangeh,
                           mapping=aes(x=Estimates, xmin=Estimates-SE, xmax=Estimates+SE),
                           stat="identity", size=0.5, colour="#696969") +
  theme_classic2()   +
  theme(legend.position = "bottom",
  panel.grid.major.x = element_line(colour="grey92", size=0.5, linetype="dashed"),
  strip.background = element_rect(fill = "grey92", 
                colour = "black", size = 1)
  )

## Collection_hist
gg.spp <- as_tibble(sample_data(ps2)) %>%
  dplyr::rename(label = psyllid_spp) %>%
  group_by(label) %>%
  summarise(n_species = n()) %>%
  left_join(p$data %>%
  filter(isTip) %>% select(c(label, y))) %>%
  filter(!is.na(y)) %>%
  ggplot(aes(x=factor(.$label, levels=.$label[order(.$y)]), y=1, fill=n_species)) +
  geom_tile() +
  coord_flip() + theme_void() +
  scale_fill_distiller(palette = "Reds", direction=1)

Fig1 + gg.spp 
  
# Change relative grid sizes
library(grid)
library(gtable)

gt = ggplot_gtable(ggplot_build(Fig1))
gt$layout$l[grep('panel-3', gt$layout$name)] # you want to find the column specific to panel-2

gt$widths[9] = 0.3*gt$widths[9] # reduce width of column 9
grid.draw(gt) # plot with grid draw
```

# Beta diversity

## PCA plot
```{r}


```



## Map of colleciton locations

```{r Collection map}
#Make spatial distance matrix
library(sp)
 # Spatial distance
  envData <- read.csv("sample_data/Sample_info.csv", header=TRUE)  %>%
    dplyr::filter(!duplicated(SampleID)) %>%
    set_rownames(.$SampleID) %>%
    separate(Collection, into= c("Lat", "Long"), sep=" ", remove = FALSE) %>%
    separate(Lat, paste("lat",c("d","m","s"), sep="_") ) %>%
    separate(Long, paste("long",c("d","m","s"), sep="_" ) ) %>%
    mutate_at(vars(-SampleID, -psyllid_spp), funs(as.numeric)) %>%
    mutate(lat=-(lat_d + lat_m/60 + lat_s/60^2),
              long=long_d + long_m/60 + long_s/60^2) %>%
    dplyr::select(SampleID, psyllid_spp, lat, long) %>%
    column_to_rownames("SampleID") %>%
    drop_na()

#Plot on map to confirm points
library("maptools")
data(wrld_simpl)
wrld_simpl@data$id <- wrld_simpl@data$NAME
wrld <- fortify(wrld_simpl, region="id")

gg.worldmap <- ggplot(envData) +
  geom_map(data=wrld, map=wrld, aes(map_id=id, x=long, y=lat), fill="lightgrey", color="#7f7f7f", size=0.25) +
  geom_point(aes(x=long, y=lat, color=psyllid_spp), alpha=.5, size=3) + 
  theme_bw() +
  coord_fixed(ylim = c(-90,90), xlim=c(-180,180))  

gg.anzmap <- ggplot(envData) +
  geom_map(data=wrld, map=wrld, aes(map_id=id, x=long, y=lat), fill="lightgrey", color="#7f7f7f", size=0.25) +
  geom_point(aes(x=long, y=lat, color=psyllid_spp), alpha=.5, size=3) + 
  theme_bw() +
  coord_fixed(ylim = c(-50,-30), xlim=c(130,180))+
  theme(legend.position = "bottom")  


gg.nzmap <- ggplot(envData) +
  geom_map(data=wrld, map=wrld, aes(map_id=id, x=long, y=lat), fill="lightgrey", color="#7f7f7f", size=0.25) +
  geom_point(aes(x=long, y=lat, color=psyllid_spp), alpha=.5, size=3) + 
  theme_bw() +
  coord_fixed(ylim = c(-50,-30), xlim=c(160,180)) 

```


```{r session-info}
# Display current R session information
sessionInfo()
```
